---
title: "Bird Count Analysis"
output: html_document
---

```{r setup, include=F}
knitr::opts_chunk$set(echo = F, comment = '')
```

## This is an awesome bird count paper
+ with bullet points
+ and more bullet points

```{r load_data, cache=T, message=F}

library(xlsx)
file_name = "HistoricalResultsByCount.xls"
sheet_name = 'HistoricalResultsByCount'
# set the index of rows to read
row_index = 232:1141
# set the column index
col_index = 6:124
count_data = read.xlsx(file_name, sheetName =sheet_name, rowIndex = row_index, colIndex = col_index)
```



```{r clean_data}
# in the species row, maps NULL species to species from previous row
# this is due to the nest table input
d = count_data
for (i in 1:(nrow(d))){
  if (is.na(d[i, 1])){
    d[i, 1] = d[i-1, 1]
  }
}
# select which columns to actually use from the
d = d[, c(1, 13, 40:116)]
# gets the colnames (uses only the year)
colnames(d) = substr(colnames(d), 2, 5)
colnames(d)[1:2] = c('species', 'val')
rownames(d) = paste(d[,1], d[,2], sepc = ':')

# coeces columns to numeric
for ( i in 3:(ncol(d))){
  d[,i] = as.numeric(d[, i])
}
# fills Nulls with zero
d[is.na(d)]= 0
```



```{r aggredate data}
# aggreate counts by species with years as columns

counts = aggregate(d[, 3:(ncol(d))], by=list(d$species), FUN=sum, d$val == ' Number:' )
# set sepcies as row names
rownames(counts) = counts[,1]
counts[,1]=NULL

# aggreate hours by species by years
hours = aggregate(d[, 3:(ncol(d))], by=list(d$species), FUN=sum, d$val == 'Num/Party Hrs.:' )
# set sepcies as row names
rownames(hours) = hours[,1]
hours[,1]=NULL

# re index the rows based on sum counts
row_index = order(apply(counts, 1, sum), decreasing = T)

counts = counts[row_index,]
hours = hours[row_index,]

```

```{r show data}
# show the top of the data set
print(head(counts)[1:10])
```





```{r load weather data}
# reads in a csv of weather data requestion from NOAA through https://www.ncdc.noaa.gov/cdo-web/confirmation
# data is a monthly average
file_name = 'chs_wx.csv'
wx_data = read.csv(file_name)
# filter to rows with only december
wx_data = wx_data[substr(wx_data$DATE,6, 8) == '12', ]
# changes the Date to use only the year, (keeping in mind this is the December average)
wx_data$DATE = as.numeric(substr(wx_data$DATE, 1, 4))

```




```

```{r noaa weather data dowload, eval=F}
# uses noaa API to download weather rirectly
library(rnoaa)
# look noaa stations at https://www.ncdc.noaa.gov/cdo-web/datatools/findstation
token = scan("noaa_api_token", what="character", sep=NULL)
datasetid = 'GHCND'
stationid = 'GHCND:US1SCCR0096'
startdate = '1939-01-01'
enddate = '2019-01-01'
wx_data = ncdc(datasetid = datasetid, stationid=stationid , startdate=startdate, enddate=enddate, token = token)


```
